---
title: "Homework 2"
author: Qianhui Xu
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
library(knitr)
```

#Problem 1

##First, define a path to the Trash Wheel dataset. 

```{r}
path_trashwheel = "./Trash-Wheel-Collection-Totals-8-6-19.xlsx"
```


##Read and clean the Mr. Trash Wheel dataset. 

```{r}
trashwheel_df = 
	read_excel(
		path = path_trashwheel,
		sheet = "Mr. Trash Wheel",
		range = cell_cols("A2:N408")) %>% 
	janitor::clean_names() %>% 
  drop_na(dumpster) %>%
	mutate(
		sports_balls = round(sports_balls),
		sports_balls = as.integer(sports_balls)
	)
```

##Read and clean precipitation data for 2018 and 2017. 

```{r}
precipitation_2017 = 
	read_excel(
	  path = path_trashwheel,
		sheet = "2017 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2017) %>% 
	relocate(year)


precipitation_2018 = 
	read_excel(
	  path = path_trashwheel,
		sheet = "2018 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2018) %>% 
	relocate(year)

```

Now combine annual precipitation dataframes. In the following code chunk, I create a `month_df` dataframe. 

```{r}
month_df = 
	tibble(
		month = 1:12,
		month_name = month.name
	)
precipitation_bind_df = 
	bind_rows(precipitation_2018, precipitation_2017)

precipitation_bind_df = 
	left_join(precipitation_bind_df, month_df, by = "month")

```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. 

There are a total of `r nrow(trashwheel_df)` rows in our final dataset.There are a total of `r nrow(precipitation_bind_df)` rows in the combined precipitation dataset of 2017 and 2018.

* The median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`
* The total precipitation in 2018 was `r precipitation_bind_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.




